\documentclass{article}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{graphicx}

\begin{document}

\title{Discretization of Common Filters}
\author{K. Loux}
\date{\today}
\maketitle

\section{Introduction}
This document contains difference equation implementations for transfer functions of common filters. Continuous-time transfer functions can be discretized using the relationship $z = e^{sT}$, where $T$ is the sampling period, $s$ is imaginary frequency and $z$ is discretized imaginary frequency.

In order for this relationship to be helpful, we need to solve for $s$, then replace all instances of $s$ in our continuous-time transfer functions with the result.  Solving for $s$, however, results in a relationship involving a natural logarithm.  We make a first-order approximation of the logarithm instead.

\begin{equation}
s = \frac{1}{T} \ln \left( z \right) \approx \frac{2}{T} \frac{1 - z^{-1}}{1 + z^{-1}}
\end{equation}

To go from a discrete-time transfer function to a difference equation, variables that are multiplied by $z^{-n}$ are replaced with the values of those variables from $n$ time steps in the past.

We use the variable $u$ to represent input and $y$ to represent output from the filters.  We give all coefficients in the form of $a_0$ through $a_n$ and $b_0$ through $b_m$ where $a$ coefficients are applied to input quantities and $b$ coefficients are applied to output quantities.  All coefficients can be scaled by $1 / b_0$, so then $b_0 = 1$ and the first $b$ coefficient required to be stored will be $b_1$.  These can be used in a difference equation as shown below.

\begin{equation}
y_k = \frac{a_0 u_k + a_1 u_{k - 1} + ... + a_n u_{k - n} - b_1 y_{k - 1} - ... - b_m y_{k - m}}{b_0}
\end{equation}

\section{First Order Low-Pass Filter}

Continuous-time transfer function:

\begin{equation}
\frac{Y}{U} = \frac{\omega}{s + \omega}
\end{equation}

\section{Second Order Low-Pass Filter}

Continuous-time transfer function:

\begin{equation}
\frac{Y}{U} = \frac{\omega^2}{s^2 + 2 \zeta \omega s + \omega^2}
\end{equation}

Difference equation coefficients:

\begin{align*}
a_0 &= \omega^2 T^2 \\
a_1 &= 2 a_0 \\
a_2 &= a_0 \\
b_0 &= 4 + 4 \zeta \omega T + \omega^2 T^2 \\
b_1 &= 2 \omega^2 T^2 - 8 \\
b_2 &= 4 - 4 \zeta \omega T + \omega^2 T^2
\end{align*}

\section{First Order High-Pass Filter}

Continuous-time transfer function:

\begin{equation}
\frac{Y}{U} = \frac{s}{s + \omega}
\end{equation}

\end{document}
